---
title: "StepsAnalysis"
author: "Asif Khan"
date: "September 6, 2017"
---

## R Markdown

The data is already nicely formatted so no need to process or change further. I will use the Data Frame StepsData for all further questions and process or transform as required per question.


```{r LoadingData}
StepsData <- read.csv("activity.csv")
```

What is the total number of steps taken per day?

Total number of steps ignoring NA values from the dataset is:
```{r StepsPerDay, echo=TRUE}
aggregate(steps ~ date, StepsData, mean, na.rm=TRUE)
```

Histogram of total number of steps taken each day:

```{r StepsHist}
hist(aggregate(steps ~ date, StepsData, sum, na.rm=TRUE )$steps, main = "Total number of steps taken each day", xlab = "Number of Steps", ylab = "Frequency")
```

The mean of the total number of steps taken each day is 
```{r meantotalsteps}
mean(aggregate(steps ~ date, StepsData, mean, na.rm=TRUE )$steps)
```

The median of the total number of steps taken each day is 
```{r mediantotalsteps} 
median(aggregate(steps ~ date, StepsData, mean, na.rm=TRUE )$steps)
```

Average Daily Activity Pattern:
```{r dailypattern}
library(data.table)
library(ggplot2)
dt <- data.table(StepsData[!is.na(StepsData$steps),])
a <- dt[, list(mean = mean(steps)), by = list(date, interval)] 
qplot
